Viscous Transport in Eroding Porous Media

### 2019-10-13 docs/transport/erosion_transport.tex
NM

\todo[inline]{The above set of references for ``examples of course-grained models" seems very thorough and perhaps not as relevant as other references on erosion, melting, etc. JFMs reference format makes them take up a lot of space. So if there are references here that you wouldn't mind dropping, it might help. I wouldn't drop all of them. I think at least 3-4 are definitely needed.}

Flow in porous media is further complicated when boundaries evolve 
dynamically in response to the fluid flow. This coupling between geometry
and flow occurs, for example, in applications involving
melting~\citep{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\citep{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\citep{joh-eli1995, hewett2018modelling},
biofilm growth~\citep{tan-val-wer2015}, and crack
formation~\citep{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\citep{berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.

@@

Owing to the scales present in groundwater flow~\citep{bea1972}, we
model the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\citep{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.  We compute the
vorticity in the fluid bulk since, on solid boundaries, vorticity
reduces to shear and thus provides a convenient way to simultaneously
visualize local erosion rates and changes in the surrounding flow
(figure~\ref{fig:Eroding50vort}).  A variety of numerical methods for
the fluid solver are available including lattice-gas or lattice
Boltzmann~\citep{kop-kat-tim1996, dar-mcc1998}, finite
volumes~\citep{suo-liu-gan2019, den-ica-hid2018,
sie-ili-pri-riv-gua2019}, finite differences~\citep{leb-ded-dav-bou2007,
knu-car2005}, and cellular automata~\citep{rot1988}.  However,
simulating erosion requires a method that is high-order, can resolve
nearly-touching arbitrarily shaped grains, and can accurately compute
derivatives of the velocity field.  Since the governing equations are
linear and homogeneous, we address these challenges by converting the
fluid equations to a boundary integral equation (BIE).

@@

To compute stable simulations of erosion, we use high-order methods in
both space and time. The time integration method is unchanged from
previous work~\citep{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be
triggered by changes in sign of the shear stress, and we use a stable
second-order Runge-Kutta method applied to the {\thL}
coordinates~\citep{hou-low-she1994} of the eroding grains.  In this
work, we introduce a new quadrature method to resolve dense suspensions.
The accuracy of the trapezoid rule, which was used in previous work, is
adequate for bodies that are sufficiently separated~\citep{tre-wei2014}.
However, it loses accuracy whenever grains are in close contact.  One of
the earliest works to address quadrature methods for nearly-singular
integrands was developed by~\citet{bak-she1986}, and in recent years,
many other near-singular integration schemes have
followed~\citep{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\citep{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

To characterize transport through the resulting configurations, particle
trajectories must be computed. Depending on the application, microscale
transport can be modelled as pure advection~\citep{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\citep{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\citep{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume trajectories to be governed by pure advection
(ie.~no diffusion), so the particle trajectories are identical to the
streamlines. We compute trajectories $\ss(t)$ that are initialized at
$\ss_0$ by solving the advection equation
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term, we solve~\eqref{eqn:tracers} with a fourth-order explicit
Runge-Kutta time stepping method.  In addition, the Barycentric
quadrature rule is used to accurately compute trajectories that are
close to an eroding grain. 
%We note that an alternative approach to
%compute trajectories is to form a distribution of velocities $\uu$ and
%use a continuous time random walk~\citep{mon-wei1965,
%den-cor-sch-ber2004, leb-den-car2008, ber-cor-den-sch2006}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\citep{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of the boundary of $[-3,3] \times [-1,1]$.  We only
place eroding bodies in $[-1,1] \times [-1,1]$ to create a buffer region
for the flow profile imposed at the inlet to transition to the more
complex flow intervening between the bodies. On each eroding body, $\nn$
is the normal vector pointing into the body, and $\ss$ is the unit
tangent vector pointing in the counterclockwise direction.  The normal
vector on $\Gamma$ is the outward unit normal.  Neglecting inertial
forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations and advancing the eroding grains.
The strength of $\UU$ is adjusted at each time step to achieve a
constant pressure drop across the channel.

@@

@@

@@

To compute the tortuosity and the anomalous dispersion rates
(section~\ref{sec:transport}), we require the streamlines that are
governed by equation~\eqref{eqn:tracers}.  To characterize the
transport, it is crucial that the streamlines are computed accurately.
If a low-order time stepping method is used, or if $\uu(\ss(t))$ is
inaccurate, then a trajectory $\ss(t)$ can unphysically enter a grain,
rendering the trajectory meaningless to quantify transport.  However, if
we simply ignore trajectories that pass close to grains, this would
significantly skew the characterization of the transport. Therefore, we
use high-order quadrature and time stepping methods.  In particular,
depending on the proximity of $\ss(t)$ to $\bd\Omega$
(section~\ref{sec:fmm}), we apply the trapezoid rule or the Barycentric
quadrature rule.  For time stepping, we use a fourth-order explicit
Runge-Kutta method.  By using these high-order methods, we are able to
simulate dynamics very close to the eroding bodies (see
figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We now present numerical results of dense packings of bodies eroding in
Stokes flow. Moreover, we analyze transport through the eroded
geometries by examining coarse-grained variables.  Each eroding body is
initially a circle with center $\cc_i$, $r_i$, and length $L_i = 2\upi
r_i$.  The center and radius are chosen at random, and the body is
accepted if it is contained in $[-1,1] \times [-1,1]$ and is
sufficiently separated from all other bodies.  Since the error of our
quadrature method is bounded independent of the geometry, we consider
bodies that are separated by less than 10\% of an arclength spacing.
The randomized method is repeated until the initial geometry reaches a
desired initial porosity.

@@

In figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For all three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop.
However, in the top half of the body, there is a sudden increase in the
shear stress, and this leads to a region of high curvature.  This
behavior is also present in three dimensions~\citep[see figure
7(c)]{mit-spa2017}.  The biggest difference between the two- and
three-dimensional results is the presence of a recirculation zone.  In
three dimensions, there is no recirculation between the solid wall and
the spherical body~\citep{cha-feu2003}, but recirculation is possible in
two dimensions~\citep{chw-wu1975, hig1985}.  To visualize the flow, we
plot the vorticity of the final time step from figure~\ref{fig:NearWall}
in figure~\ref{fig:NearWall_vort}.  In these examples, a small
recirculation zone, both in size and magnitude, is present in the region
where the vorticity is smallest

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding grains with the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
imposed on $\Gamma$. The flow rate $U$ is chosen so that the average
pressure drop from $x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$
once all the grains have vanished.  The vorticity and grain
configuration at four equispaced times are shown in
figure~\ref{fig:Eroding20vort}.  Initially, several of the grains are
closer to the outer wall than the $5h$ threshold required for the
trapezoid rule to achieve machine precision.  In particular, the
distance between bodies 1, 6, 13, and 15 and the outer wall is $1.3h$,
$2.9h$, $2.8h$, and $1.3h$, respectively, where $h$ is the arclength
spacing of the outer wall $\Gamma$.  In addition, the distance between
several pairs of eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and
14 \& 18, is too small to be accurately resolved with the trapezoid
rule.  By using the Barycentric quadrature rule, the interaction between
these nearly-touching bodies is accurately resolved to the desired
accuracy, and erosion can be simulated until all the bodies have
vanished.

@@

We next analyze the effect of erosion on the area fraction and the flow
rate.  In figure~\ref{fig:Eroding20flowrate}(a), we plot the area
fraction as a function of normalized time.  The general trend of the
area fraction resembles our previous work~\citep[see][figure
10(a)]{qua-moo2018}, but with a larger initial area fraction.  In
figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\citep[see][figure 10(b)]{qua-moo2018}, except that the initial
flow rate is an order of magnitude smaller because of the larger initial
area fraction.  Starting around normalized time $0.2$,
figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as an exponential law.  The line of
best fit is
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in figure~\ref{fig:Eroding20flowrate}(b).

@@

In figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity. The initial porosity is $\phi = 37.68\%$, and
the initial tortuosity is $T = 1.16$.  The tortuosity is computed with
both the length of the streamlines~\eqref{eqn:tortuosity1} (red stars)
and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and from applying
quadrature to compute the tortuosity.  As the bodies erode, wide
channels form where streamlines undergo only minor vertical deflections,
and this explains why the tortuosity eventually decreases with porosity.
We computed lines of best fit using the porosity-tortuosity
models~\eqref{eqn:tortuosityModels} and found that the power law
minimizes the error.  The  black dashed line in
figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064},
\end{align}
with a root-mean-square error of $5.90 \times 10^{-3}$.  Interestingly,
at the low porosities, the tortuosity initially increases. This increase
occurs because in the absence of erosion (left plot in
figure~\ref{fig:Eroding20vort}), many of the streamlines, such as those
initialized between bodies 15 \& 18, only perform minor deflections to
pass through the narrow regions, albeit, very slowly.  However, as
erosion starts to open the channels, the streamlines deflect into the
fast regions, such as the region above body 11, and this increases the
amount of vertical deflection, and therefore the tortuosity. While this
increase in tortuosity is interesting, in the next two examples we will
see that the tortuosity does not initially increase.

@@

We compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2} at each time step. The initial porosity
is $\phi = 30.67\%$ and the initial tortuosity is $T = 1.24$.  In
figure~\ref{fig:ErodingLow20Transport}(a), we plot the tortuosity with
respect to the porosity (blue) and the line of best fit (black) using
the power law 
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.1669}.
\end{align}
This model outperforms the other three models in
equation~\eqref{eqn:tortuosityModels}, and its root-mean-squared error
is $1.13 \times 10^{-2}$.  As grains erode, there is an increase in the
number of streamlines that take a nearly direct path through the
geometry, and this decreases the tortuosity. However, the channelization
effect of erosion results in an increase in the tortuosity since the
length of many of the streamlines increases when they deflect from a
high porosity region (low pressure) to a low porosity region (high
pressure). For this example, we see that the net effect is a decrease in
the tortuosity for all time.

@@

@@

The transport is characterized in terms of tortuosity and anomalous
dispersion. While the local tortuosity agrees qualitatively with other
works~\citep{mat-kha-koz2008}, the tortuosity of eroded geometries
cannot be completely described in terms of the porosity. In particular,
we observe that for certain porous media, the tortuosity can increase as
erosion increases the porosity of the porous media. We also observe
super-dispersive spreading, and the rate of dispersion significantly
depends not only on the porosity, but also the number of eroding bodies
and their distribution.

To further our understanding of erosion, we are examining other bulk and
statistical properties of an eroding porous media. In this work, we
provide preliminary results for the pore throat sizes which affect the
anomalous dispersion rate~\citep{dea-qua-bir-jua2018}.  At a later date,
we will report results on the  development of anisotropic effects and
the distributions of grain sizes, shapes, and opening angles.

>>>>

\todo[inline]{The above set of references for ``examples of course-grained models" seems very thorough and perhaps not as relevant as other references on erosion, melting, etc. JFM's reference format makes them take up a lot of space. So if there are references here that you wouldn't mind dropping, it might help. I wouldn't drop all of them. I think at least 3-4 are definitely needed.}

Flow in porous media is further complicated when boundaries evolve 
dynamically in response to the fluid flow. This coupling between geometry
and flow occurs, for example, in applications involving
melting~\citep{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\citep{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\citep{joh-eli1995, hewett2018modelling},
biofilm growth~\citep{tan-val-wer2015}, and crack
formation~\citep{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\citep{ris-moo-chi-she-zha2012, berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.

@@

Owing to the scales present in groundwater flow~\citep{bea1972}, we
model the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\citep{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.  We compute the
vorticity in the fluid bulk since, on solid boundaries, vorticity
reduces to shear and thus provides a convenient way to simultaneously
visualize local erosion rates and changes in the surrounding flow
(figure~\ref{fig:Eroding50vort}).  A variety of numerical methods for
the fluid solver are available including lattice-gas or lattice
Boltzmann~\citep{kop-kat-tim1996, dar-mcc1998}, finite
volumes~\citep{suo-liu-gan2019, den-ica-hid2018,
sie-ili-pri-riv-gua2019}, finite differences~\citep{leb-ded-dav-bou2007,
knu-car2005}, and cellular automata~\citep{rot1988}.  
\todo[inline]{The set of references above seems to me like an unneeded `straw-man' argument. i.e.~Just because we are using a boundary integral method, does that mean we have to list every other possible way of solving the fluid equations? It takes up a fair amount of space. On the other hand, if there are references here you want to keep, by all means do it.}
However, simulating erosion requires a method that is high-order, can resolve
nearly-touching arbitrarily shaped grains, and can accurately compute
derivatives of the velocity field.  Since the governing equations are
linear and homogeneous, we address these challenges by converting the
fluid equations to a boundary integral equation (BIE).

@@

To compute stable simulations of erosion, we use 
\nick{methods of high order in both space and time.}
The time integration is unchanged from
previous work~\citep{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities that can be
triggered by changes in sign of the shear stress, and we use a stable
second-order Runge-Kutta method applied to the {\thL}
coordinates~\citep{hou-low-she1994} of the eroding grains.  In this
work, we introduce a new quadrature method to resolve dense suspensions.
The accuracy of the trapezoid rule, which was used in previous work, is
adequate for bodies that are sufficiently separated~\citep{tre-wei2014}, 
but not for grains in close contact.  
One of the earliest quadrature methods for nearly-singular
integrands was developed by~\citet{bak-she1986}, and in recent years,
many other \nick{\st{near-singular integration}} schemes have
followed~\citep{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
method~\citep{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method to compute
the velocity gradient, which is needed to evaluate the shear stress and
the fluid vorticity.

To characterize transport through the resulting configurations, particle
trajectories must be computed. Depending on the application, microscale
transport can be modelled as pure advection~\citep{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\citep{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\citep{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume trajectories to be governed by pure advection
(ie.~no diffusion), so the particle trajectories are identical to the
streamlines. We compute trajectories $\ss(t)$ that are initialized at
$\ss_0$ by solving the advection equation
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
\todo[inline]{Could the above equation be changed to inline to save space?}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term, we solve~\eqref{eqn:tracers} with a fourth-order explicit
Runge-Kutta time stepping method. The Barycentric
quadrature rule is used to accurately compute trajectories that are
close to an eroding grain. 
%We note that an alternative approach to
%compute trajectories is to form a distribution of velocities $\uu$ and
%use a continuous time random walk~\citep{mon-wei1965,
%den-cor-sch-ber2004, leb-den-car2008, ber-cor-den-sch2006}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\citep{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of the boundary of $[-3,3] \times [-1,1]$.  We only
place eroding bodies in $[-1,1] \times [-1,1]$ to create a buffer region
that allows the flow profile imposed at the inlet to transition to the more
complex flow intervening between the bodies. On each eroding body, $\nn$
is the normal vector pointing into the body, and $\ss$ is the unit
tangent vector pointing in the counterclockwise direction.  The normal
vector on $\Gamma$ is the outward unit normal.  Neglecting inertial
forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations and advancing the eroding grains.
The strength of $\UU$ is adjusted at each time step to achieve a
constant pressure drop across the channel, 
\nick{motivated by the geological situation of a porous medium connecting two regions of  fixed hydraulic heads.}

@@

\todo[inline]{The above vertical fraction could be made into a horizontal fraction, which would save a lot of space.}

@@

\todo[inline]{More prime examples of vertical fractions that could be made horizontal.}

@@

To examine the tortuosity and the anomalous dispersion rates
(section~\ref{sec:transport}), we require accurate computation of the streamlines
governed by equation~\eqref{eqn:tracers}.
\nick{\st{To characterize the transport, it is crucial that the streamlines are computed accurately.}}
If a low-order time stepping method is used, or if $\uu(\ss(t))$ is
inaccurate, then a trajectory $\ss(t)$ can unphysically enter a grain,
rendering the trajectory meaningless. \nick{\st{to quantify transport.}}
\nick{\st{However, if we simply ignore trajectories that pass close to grains, this would
significantly skew the characterization of the transport. } 
However, simply ignoring trajectories that pass close to a grain could significantly bias the characterization of transport.}
Therefore, we use high-order quadrature and time stepping methods.  In particular,
depending on the proximity of $\ss(t)$ to $\bd\Omega$
(section~\ref{sec:fmm}), we apply the trapezoid rule or the Barycentric
quadrature rule.  For time stepping, we use a fourth-order explicit
Runge-Kutta method.  By using these high-order methods, we are able to
simulate dynamics very close to the eroding bodies (see
figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We now present numerical results of dense grain packings eroding in
Stokes flow and analyze transport through the evolving
geometries. \nick{\st{by examining coarse-grained variables}}.  
Each \nick{\st{eroding}} body is
initialized as a circle of center $\cc_i$, radius $r_i$, and length $L_i = 2\upi r_i$.  
\todo[inline]{There were too many sequential uses of the word `eroding' above, so I cut two of them out.}
The center and radius are chosen at random, and the body is
accepted if it is contained in $[-1,1] \times [-1,1]$ and is
sufficiently separated from all other bodies.  
\awk{Since the error of our quadrature method is bounded independent of the geometry, we consider bodies that are separated by less than 10\% of an arclength spacing.}
\todo[inline]{The above sentence seemed like a non sequitur to me. See if I got the right idea across in the sentence below.}
\nick{Owing to our adaptive quadrature rule, we can consider bodies that are separated by less than 10\% of an arclength spacing.}
The randomized method is repeated until the initial geometry reaches a
desired initial porosity.

@@

In figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For all three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop.
However, in the top half of the body, there is a sudden increase in the
shear stress, and this leads to a region of high curvature.  This
behavior is also present in three dimensions~\citep[see figure
7(c)]{mit-spa2017}.  The biggest difference between the two- and
three-dimensional results is the presence of a recirculation zone.  In
three dimensions, there is no recirculation between the solid wall and
the spherical body~\citep{cha-feu2003}, but recirculation is possible in
two dimensions~\citep{chw-wu1975, hig1985}.  To visualize the flow, we
plot the vorticity of the final time step from figure~\ref{fig:NearWall}
in figure~\ref{fig:NearWall_vort}.  In these examples, a small
recirculation zone, both in size and magnitude, is present in the region
where the vorticity is \nick{smallest.}
\todo[inline]{I was proud of myself for catching the missing period at the end, so had to point it out :)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding grains with the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
imposed on $\Gamma$. The flow rate $U$ is chosen so that the average
pressure drop from $x=-2$ to $x=2$ is held fixed at 8. Therefore, $U=1$
once all the grains have vanished.  The vorticity and grain
configuration at four equispaced times are shown in
figure~\ref{fig:Eroding20vort}.  Initially, several of the grains are
closer to the outer wall than the $5h$ threshold required for the
trapezoid rule to achieve machine precision.  In particular, the
distance between bodies 1, 6, 13, and 15 and the outer wall is $1.3h$,
$2.9h$, $2.8h$, and $1.3h$, respectively, where $h$ is the arclength
spacing of the outer wall $\Gamma$.  In addition, the distance between
several pairs of eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and
14 \& 18, is too small to be accurately resolved with the trapezoid
rule.  By using the Barycentric quadrature rule, the interaction between
these nearly-touching bodies is accurately resolved to the desired
accuracy, and erosion can be simulated until all the bodies have
vanished.
\todo[inline]{accurately resolved to the desired accuracy is redundant.}

@@

We next analyze the effect of erosion on the area fraction and the flow
rate.  In figure~\ref{fig:Eroding20flowrate}(a), we plot the area
fraction as a function of normalized time.  The general trend of the
area fraction resembles our previous work~\citep[see][figure
10(a)]{qua-moo2018}, but with a larger initial area fraction.  In
figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\citep[see][figure 10(b)]{qua-moo2018}, except that the initial
flow rate is an order of magnitude smaller because of the larger initial
area fraction.  Starting around normalized time $0.2$,
figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as an exponential law.  The line of
best fit is
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in figure~\ref{fig:Eroding20flowrate}(b).
\todo[inline]{Perhaps equations like the above could be made inline to save space.}

@@

In figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity. The initial porosity is $\phi = 37.68\%$, and
the initial tortuosity is $T = 1.16$.  The tortuosity is computed with
both the length of the streamlines~\eqref{eqn:tortuosity1} (red stars)
and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and from applying
quadrature to compute the tortuosity.  As the bodies erode, wide
channels form where streamlines undergo only minor vertical deflections,
and this explains why the tortuosity eventually decreases with porosity.
We computed lines of best fit using the porosity-tortuosity
models~\eqref{eqn:tortuosityModels} and found that the power law
minimizes the error.  The  black dashed line in
figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064},
\end{align}
\todo[inline]{This one definitely seems like a candidate for an inline equation.}
with a root-mean-square error of $5.90 \times 10^{-3}$.  Interestingly,
at the low porosities, the tortuosity initially increases. This increase
occurs because in the absence of erosion (left plot in
figure~\ref{fig:Eroding20vort}), many of the streamlines, such as those
initialized between bodies 15 \& 18, only perform minor deflections to
pass through the narrow regions, albeit, very slowly.  However, as
erosion starts to open the channels, the streamlines deflect into the
fast regions, such as the region above body 11, and this increases the
amount of vertical deflection, and therefore the tortuosity. While this
increase in tortuosity is interesting, in the next two examples we will
see that the tortuosity does not initially increase.

@@

We compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2} at each time step. The initial porosity
is $\phi = 30.67\%$ and the initial tortuosity is $T = 1.24$.  In
figure~\ref{fig:ErodingLow20Transport}(a), we plot the tortuosity with
respect to the porosity (blue) and the line of best fit (black) using
the power law 
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.1669}.
\end{align}
\todo[inline]{inline?}
This model outperforms the other three models in
equation~\eqref{eqn:tortuosityModels}, and its root-mean-squared error
is $1.13 \times 10^{-2}$.  As grains erode, there is an increase in the
number of streamlines that take a nearly direct path through the
geometry, and this decreases the tortuosity. However, the channelization
effect of erosion results in an increase in the tortuosity since the
length of many of the streamlines increases when they deflect from a
high porosity region (low pressure) to a low porosity region (high
pressure). For this example, we see that the net effect is a decrease in
the tortuosity for all time.

@@

\todo[inline]{Very long captions in Figs. 10 and 12 and maybe others.}

@@

The transport is characterized in terms of tortuosity and anomalous
dispersion. While the local tortuosity agrees qualitatively with other
works~\citep{mat-kha-koz2008}, the tortuosity of eroded geometries
cannot be completely described in terms of the porosity. In particular,
we observe that for certain \nick{\st{porous media} configurations}, the tortuosity \nick{transiently increases, even though the porosity always increases due to erosion.}
\nick{\st{as erosion increases the porosity of the porous media.}} 
We also observe super-dispersive spreading, and the rate of dispersion significantly
depends not only on the porosity, but also the number of eroding bodies
and their distribution.

To further our understanding of erosion, we are examining other bulk and
statistical properties of an eroding porous media. In this work, we
provide \nick{\st{preliminary}} results for the pore throat sizes which affect the
anomalous dispersion rate~\citep{dea-qua-bir-jua2018}.  At a later date,
we will report results on the  development of anisotropic effects and
the distributions of grain sizes, shapes, and opening angles.
